Examining the effect of a trend on R*

### 2020-03-21 src/s_trends.Rmd
added sims of trends

for (i in 1:nrow(conds)) {
  iters <- conds[i, "iters"]
  trend <- conds[i, "trend"]
  rep <- conds[i, "rep"]
  dimension <- conds[i, "dimension"]
  r <- array(rnorm(iters * chains * dimension), c(iters, chains, dimension))
  r[, , 1] <- r[, , 1] + seq(-trend, trend, length.out = iters)
  rs <- as.data.frame(monitor_extra(r))
  r_2d <- matrix(nrow = iters * chains, ncol=(dimension + 2))
  k <- 1
  for(i in 1:iters){
    for(j in 1:chains){
      r_2d[k, ] <- c(i, j, r[i, j, ])
      k <- k + 1
    }
  }
  r_2d <- r_2d %>%
    as.data.frame()
  colnames(r_2d)[1:2] <- c("iter", "chain")

@@

@@

conds <- expand.grid(
  iters = c(250, 1000, 4000), 
  trend = c(0, 0.25, 0.5, 0.75, 1),
  rep = 1:10,
  rho = c(0, 0.25, 0.5, 0.75, 1)
)
res <- vector("list", nrow(conds))
chains = 4
sigma <- 1

for (i in 1:nrow(conds)) {
  iters <- conds[i, "iters"]
  trend <- conds[i, "trend"]
  rep <- conds[i, "rep"]
  rho <- conds[i, "rho"]
  r <- array(rnorm(iters * chains), c(iters, chains))
  for(i in 1:chains)
    r[, i] <- f_ar1(rho, sigma, iters)
  rs <- as.data.frame(monitor_extra(r))

  # R* calculation
  r1 <- r %>% 
    as.data.frame() %>% 
    melt(id.vars = NULL) %>% 
    mutate(chain=as.factor(variable)) %>% 
    select(-variable)
  if(dimension==1)
    r1 <- r1 %>% mutate(ones=rnorm(length(chain)))
  a_accuracy <- f_fit_gbm_predict(r1)

  # Split R* calculation
  half_rows <- nrow(r) / 2
  r2 <- cbind(r[1:half_rows, ], r[(half_rows+1):nrow(r), ]) %>% 
    as.data.frame() %>% 
    melt(id.vars = NULL) %>% 
    mutate(chain=as.factor(variable)) %>% 
    select(-variable)
  if(dimension==1)
    r2 <- r2 %>% mutate(ones=rnorm(length(chain)))
  b_accuracy <- f_fit_gbm_predict(r2)

  res[[i]] <- cbind(iters, trend, rep, rho, rs, "Rstar"=a_accuracy * n_distinct(r1$chain),
                    "Rstar_split"=b_accuracy * n_distinct(r2$chain))
}
res <- bind_rows(res)
saveRDS(res, "../output/experiment_trends_rho.rds")

@@

>>>>

for (i in 1:nrow(conds)) {
  iters <- conds[i, "iters"]
  trend <- conds[i, "trend"]
  rep <- conds[i, "rep"]
  dimension <- conds[i, "dimension"]
  r <- array(rnorm(iters * chains * dimension), c(iters, chains, dimension))
  r[, , 1] <- r[, , 1] + seq(-trend, trend, length.out = iters)
  rs <- as.data.frame(monitor_extra(r))
  r_2d <- matrix(nrow = iters * chains, ncol=(dimension + 2))
  k <- 1
  for(ii in 1:iters){
    for(j in 1:chains){
      r_2d[k, ] <- c(ii, j, r[ii, j, ])
      k <- k + 1
    }
  }
  r_2d <- r_2d %>%
    as.data.frame()
  colnames(r_2d)[1:2] <- c("iter", "chain")

@@

```{r}
res <- readRDS("../output/experiment_trends_one_dim.rds") %>% 
  select(trend, iters, rep, dimension, Rstar) %>% 
  unique()
ggplot(data = res, aes(x = trend, y = Rstar)) + 
  geom_point() + 
  geom_jitter() + 
  facet_grid(vars(iters), vars(dimension)) + 
  geom_hline(yintercept = 1) + 
  ggtitle('R* without splitting')
```

```{r}
res <- readRDS("../output/experiment_trends_one_dim.rds") %>% 
  select(trend, iters, rep, dimension, Rstar_split) %>% 
  unique()
ggplot(data = res, aes(x = trend, y = Rstar_split)) + 
  geom_point() + 
  geom_jitter() + 
  facet_grid(vars(iters), vars(dimension)) + 
  geom_hline(yintercept = 1) + 
  ggtitle('Split-R*')
```

@@

conds <- expand.grid(
  iters = c(250, 1000, 4000), 
  trend = c(0, 0.25, 0.5, 0.75, 1),
  rep = 1:10,
  rho = c(0.8, 0.85, 0.9, 0.95, 0.975, 1)
)
res <- vector("list", nrow(conds))
chains = 4
sigma <- 1

for (i in 1:nrow(conds)) {
  iters <- conds[i, "iters"]
  trend <- conds[i, "trend"]
  rep <- conds[i, "rep"]
  rho <- conds[i, "rho"]
  r <- array(rnorm(iters * chains), c(iters, chains))
  for(j in 1:chains)
    r[, j] <- f_ar1(rho, sigma, iters)
  rs <- as.data.frame(monitor_extra(r))

  # R* calculation
  r1 <- r %>% 
    as.data.frame() %>% 
    melt(id.vars = NULL) %>% 
    mutate(chain=as.factor(variable)) %>% 
    select(-variable)
  if(ncol(r1)==2)
    r1 <- r1 %>% mutate(ones=rnorm(length(chain)))
  a_accuracy <- f_fit_gbm_predict(r1)

  # Split R* calculation
  half_rows <- nrow(r) / 2
  r2 <- cbind(r[1:half_rows, ], r[(half_rows+1):nrow(r), ]) %>% 
    as.data.frame() %>% 
    melt(id.vars = NULL) %>% 
    mutate(chain=as.factor(variable)) %>% 
    select(-variable)
  if(ncol(r2)==2)
    r2 <- r2 %>% mutate(ones=rnorm(length(chain)))
  b_accuracy <- f_fit_gbm_predict(r2)

  res[[i]] <- cbind(iters, trend, rep, rho, rs, "Rstar"=a_accuracy * n_distinct(r1$chain),
                    "Rstar_split"=b_accuracy * n_distinct(r2$chain))
}
res <- bind_rows(res)
saveRDS(res, "../output/experiment_trends_rho_nearer1.rds")

@@

```{r}
res <- readRDS("../output/experiment_trends_rho_nearer1.rds")

ggplot(data = res, aes(x = rho, y = Rstar)) + 
  geom_jitter() + 
  facet_grid(vars(iters)) + 
  geom_hline(yintercept = 1) + 
  ggtitle('R* without splitting') +
  ylim(0, 6)

ggplot(data = res, aes(x = rho, y = Rstar_split)) + 
  geom_jitter() + 
  facet_grid(vars(iters)) + 
  geom_hline(yintercept = 1) + 
  ggtitle('Split-R*') +
  ylim(0, 6)
```
